---
title: "Plotting traced Lateral Horn Neurons"
output: rmarkdown::html_vignette
author: "Gregory Jefferis, Johannes Kohl, Aaron Ostrovsky, Shahar Frechter"
vignette: >
  %\VignetteIndexEntry{Plotting traced Lateral Horn Neurons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Intro
This vignette shows some examples of how to use neuronal tracings
in combination with GJ's [nat](https://github.com/jefferis/nat)
(NeuroAnatomy Toolbox) R package to analyse
the structure of lateral horn neurons in this study.

## Installation

```{r setup, include=FALSE}
library(frulhns)
# note setting fig.path like this (i.e. putting figures in the same folder as the other files)
# works around a subtle bug in knitr where the 'figure' subfolder is not created by a custom plot
# hook such as hook_rgl (but this folder would be created by a regular 'plot').
knitr::opts_chunk$set(
  fig.height = 4,
  fig.width = 6,
  collapse = TRUE,
  comment = "#>"
)
options(replace.assign=TRUE,width=55)
# so that we can make snapshots from rgl plots
rgl::setupKnitr()
tryCatch({
  library(nat)
},error = function(e) {
  stop("Please install nat package from https://github.com/jefferis/nat")
})
```

Before running this vignette, it is necessary to install an R package
nat (NeuroAnatomy Toolbox). This will have happened if you installed
the frulhns package. There are instructions at \url{https://github.com/jefferis/nat},
but the quickest method is to run the following code in an interactive
session:

```{r install-instructions, eval=FALSE}
install.packages("nat")
```

## Basic Plots

Now we can start a fresh R session, load required libraries and get
to work

```{r eval=FALSE}
library(frulhns)
library(nat)
```

Let's start by plotting a single neuron before registration

```{r single-neuron-raw, rgl=TRUE}
plot3d(jknraw[[1]])
```

First 10 neurons

```{r first-10-neurons-raw, rgl=TRUE}
plot3d(jknraw[1:10], col = rainbow, WithNodes = FALSE)
```

Now looking at multiple neurons before registration doesn't make much
sense, so let's try again using the registered neurons

```{r first-10-neurons, rgl=TRUE}
plot3d(jkn[1:10], col = rainbow, WithNodes = FALSE)
```

## Brains and Views

We can already see that those neurons are more tightly distributed.
However it's not really too clear how they relate to the structure
of the brain. Let's define a quick function to plot a brain surface.

```{r IS2Surf}
#' Use rgl to plot a 3d surface of the brain
#' 
#' @param alpha Alpha value for surface
#' @param col Colour of surface
#' @param ... Additional parameters passed to \code{plot3dsurface}
#' @export
#' @examples
#' @seealso \code{plot3dsurface} in \code{AnalysisSuite} code bundle.
#' is2surf()
#' is2surf(alpha=0.4,col='cyan')
is2surf <- function(alpha = 0.2, col = 'grey', ...) {
  plot3d(IS2Surf, "Exterior", col = col, alpha = alpha, ...)
}
```

Now we can use that to provide some context:

```{r first-10-neurons-brain, rgl=TRUE}
clear3d()
is2surf()
plot3d(jkn[1:10], col = rainbow, WithNodes = FALSE)
```

The default view is a actually a dorsal one, that is looking down
the dorso-ventral axis onto the dorsal surface of the brain. People
are actually more used to looking at a frontal view, but we can use
the \texttt{nview3d} utility function to help plot the same data in
a more conventional view:

```{r first-10-neurons-brain-frontal, rgl=TRUE}
is2surf()
plot3d(jkn[1:10], col = rainbow, WithNodes = FALSE)
nview3d("frontal", zoom=0.7)
```

```{r first-10-neurons-brain-frontal-zoom, rgl=TRUE}
is2surf()
plot3d(jkn[1:10], col = rainbow, WithNodes = FALSE)
# Hmm, let's zoom in a bit, notice that zoom values
# get smaller as you zoom in (see ?nview3d for details)
nview3d("frontal", zoom = 0.5)
```

## Plotting selected neurons

OK, now let's look at a particular set of lateral horn neurons: wild
type aSP-g. We can use the \texttt{dataframe} attached to the \texttt{neuronlist}
to select which neurons to plot and how to colour them.

```{r asgp-wildtype, rgl=TRUE}
plot3d(jkn,
       subset=cluster=="aSP-g" & shortGenotype=="JK1029",
       col=sex, colpal=c(male='cyan',female='magenta'), 
       WithNodes=FALSE,lwd=2)
nview3d("frontal", zoom=0.7)
```

The distinction between male and female neurons is clear with several
regions that are exclusively innervated by male or female arbours.

Similarly let's look at all the neurons recored from the wildtype
JK1029-Gal4 male flies.

```{r JK1029-wildtype-male, rgl=TRUE}
plotresult=plot3d(jkn, subset=shortGenotype=="JK1029" & sex=="male",
                  col=cluster,WithNodes=FALSE,lwd=2)
nview3d("frontal", zoom=0.7)
# note that plotresult includes a useful dataframe as an attribute
# with details about the selected neurons
head(attr(plotresult,'df'))
```

There are 3 clusters. The cell bodies for aSP-f (red) are more dorsal
whereas those for aSP-g and aSP-h (green, blue) are more lateral and
not obviously separable.
